---
title: "mhAnalysis"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{MPO_vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```



## Authors
Erqiang Hu: Department of Bioinformatics, School of Basic Medical
Sciences, Southern Medical University.

Shanye Yin： Albert Einstein College of Medicine


```{r setup}
library(mhAnalysis)
```

## Introduction
A significant proportion of human cancers result from the integration of viruses
into the human genome. Therefore, investigating the mechanisms of viral 
integration is crucial for understanding the underlying mechanisms of many 
devastating diseases. Microhomology (MH) analysis at viral integration sites 
plays a critical role in the study of viral integration mechanisms. 
Previous research has shown a significant enrichment of microhomology (MH) 
between the human genome and the human papillomavirus (HPV) genome at 
integration breakpoints or nearby regions. To facilitate MH analysis, we have 
developed the mhAnalysis software package. The following sections provide an 
overview of the principles and usage of mhAnalysis.



### (1) Integration sites Information:
SurVirus is an excellent software tool for analyzing viral integration sites 
with low false-positive and false-negative rates, providing researchers with 
reliable data. Therefore, mhAnalysis performs microhomology analysis by 
integrating with the results obtained from SurVirus.

### (2)Obtaining Integrated Sequences and Host-Virus Sequences at Integration sites:
In MH analysis, it is essential to obtain the integrated sequences, host 
sequences, and viral sequences at the integration sites. The retrieval of 
integrated sequences depends on the coordinates of the host and viral 
integration sites as well as their positive and negative strands. However, in 
some cases, SurVirus results may misidentify positive and negative strands. 
For example, if one integrated sequence has read1 recognizing the host's 
positive strand and read2 recognizing the virus's negative strand, SurVirus 
might incorrectly conclude that the host's positive strand integrated with the 
virus's negative strand when, in reality, the host's positive strand integrated 
with the virus's positive strand. To address this, we provide the add_strand() 
function to add accurate positive and negative strand information to the 
integrated sequences. This function first employs the internal get_r() function 
to identify which end each sequence belongs to and then selects single-end 
sequences that match both the virus and the host, using this information to 
determine the positive and negative strands for the virus and host.

When there is MH at the insertion sites, SurVirus results may suffer from 
misalignment issues. The misalignment length corresponds to the microhomology 
length at the insertion site. To rectify this and avoid complications in 
microhomology length statistics, we offer the get_real_loc() function, which 
identifies the genuine integration coordinates.

### (3)Analyzing MH Within Given Flanking Regions:
We provide the add_mh_flank() function to append MH length information within 
specified flanking regions to the SurVirus results. The process involves using 
the get_seq() function to add host sequence and virus sequence information 
within the defined flanking regions. These two sequences are combined to create 
the integrated sequence data. Subsequently, the virus sequence, host sequence, 
and integrated sequence are used as inputs to compute MH lengths using the 
get_mh function. The microhomology information is then incorporated into the 
SurVirus results.
In conclusion, we have developed the mhAnalysis package to collaborate with the 
SurVirus software, particularly for MH analysis. Through mhAnalysis, we can 
accurately address MH issues at insertion sites, ensuring the accuracy and 
reliability of insertion site data. The introduction of this tool not only 
facilitates in-depth virology research but also provides powerful analytical 
tools in the field of bioinformatics to support the exploration of disease 
mechanisms and biological interactions.

##  :hammer: Installation
```{r eval=FALSE}
devtools::install_github("huerqiang/mhAnalysis")
```

## Example

```{r eval=FALSE}
SurVirus_dir <- "SurVirus_result"
bam_dir <- "SurVirus_result/readsx"

result_t1 <- fread(file.path(SurVirus_dir, "results.t1.txt"), sep = " ", header = FALSE, fill = TRUE)
results <- fread(file.path(SurVirus_dir, "results.txt"), sep = " ", header = FALSE, fill = TRUE)
class(results) <- class(result_t1) <- "data.frame"
rownames(results) <- paste("ID", results[, 1], sep = "=")
results <- results[result_t1[, 1], ]
result_rel <- add_strand(results)
result_rel <- get_real_loc(result_rel)
result_rel <- get_seq(result_rel)
result_rel <- add_mh_flank(result_rel, len = 5)
```


```{r}
sessionInfo()
```



